{\tt APFELgrid}: a fast interface to hadronic observables for fits of parton densities

### 2016-05-04 paper.tex
Abstract and intro edits

\begin{abstract}
  We present a new software package designed to aid the inclusion of
  hadronic observables into Parton Distribution Function (PDF)
  fits. The {\tt APFELgrid} package converts interpolated weight
  tables provided by {\tt APPLgrid} files into a more efficient format
  for PDF fitting by the pre-computation and combination with PDF and
  $\alpha_s$ evolution factors provided by {\tt APFEL}. This
  combination significantly reduces the number of operations required
  to perform the calculation of hadronic observables in PDF fits and
  simplifies the structure of the calculation into a readily optimised
  scalar product. We demonstrate that our technique can lead to a
  substantial speed improvement as compared to the existing interfaces
  without any reduction in numerical accuracy.
  % We present a new method conceived to ease the inclusion of
  % hadronic observables into parton distribution function (PDF)
  % fits. Our technique relies on the the same principles adopted by
  % the existing fast interfaces but implements a set of improvements
  % aimed at optimizing the performance in the context of PDF fits
  % where many iterations and many predictions for each iteration are
  % usually required. The main novelties are the precomputation of the
  % PDF and $\alpha_s$ evolution and the optimisation of the numerical
  % convolution between hard cross sections and PDFs. We demonstrate
  % that our technique can lead to a substantial speed improvement as
  % compared to the existing interfaces.
\end{abstract}

@@

The ability to perform fast and accurate cross-section predictions for
measurements at the Large Hadron Collider (LHC) is a vital requirement
for the most precise PDF determinations. LHC measurements have a
unique capacity to shed light upon the dynamics of the proton and
provide stringent constraints upon the proton PDF. In order to make
full use of current and future LHC data, efficient strategies for the
computation of predictions for these observables must be
deployed. Software packages performing predictions for non-inclusive
hadron-collider observables cannot be easily deployed in a PDF fit
requiring rapid iteration due the time required to obtain accurate
results (usually of the order of a few hours or more per data
point). In order to overcome this limitation, the typical strategy
adopted for fast cross-section prediction relies on the precomputation
of the partonic hard cross sections in such a way that the numerical
convolution with any set of PDFs can be quickly performed by means of
interpolation techniques.

This interpolation strategy is implemented in the {\tt
  APPLgrid}~\cite{Carli:2010rw} and {\tt
  FastNLO}~\cite{Wobisch:2011ij} projects. For the computation of the
hard cross-sections, these packages rely on external codes to which
they are interfaced by means of a suite of functions allowing for the
filling of PDF- and $\alpha_s$-independent look-up tables of
cross-section weights. Codes such as {\tt MCFM}~\cite{Campbell:2010ff}
and {\tt NLOJet++}~\cite{Nagy:2003tz} have been interfaced directly to
{\tt APPLgrid}/{\tt FastNLO} and more recently de\-di\-ca\-ted
interfaces to automated general-purpose Monte Carlo (MC) event
generators have been developed. The {\tt
  aMCfast}~\cite{Bertone:2014zva} and {\tt
  MCgrid}~\cite{DelDebbio:2013kxa} codes have been specifically
designed to extract the relevant information from the {\tt
  MadGraph5\_aMC@NLO}~\cite{Alwall:2014hca} and {\tt
  SHERPA}~\cite{Gleisberg:2008ta} event generators respectively.

While these tools have proven to be extremely useful in the extraction
of parton densities, the volume of experimental data made available by
LHC collaborations for use in PDF fits is already stretching the
capabilities of the typical fitting technology. Presently a typical
global PDF fit includes thousands of hadronic data points for which
predictions have to be computed thousands of times during the
minimisation process. As a consequence the ``standard'' tools, $i.e.$
{\tt APPLgrid} and {\tt FastNLO}, might not be fast enough to meet the
needs of modern PDF fitters. For this reason a high-performance tool
tailored specifically for PDF analyses is becoming increasingly
important.

The {\tt FastKernel} method, developed in the framework of the NNPDF
global analyses~\cite{Ball:2010de,Ball:2014uwa}, was developed to
address this problem. This method differs from the standard procedure
\`{a} la {\tt APPLgrid} or {\tt FastNLO} in that it maximises the
amount of information that is precomputed prior to fitting such as to
minimise the amount of operations to be performed during the
fit. Specifically, the {\tt FastKernel} method relies on the
precomputation of the partonic hard cross-sections and DGLAP evolution
kernels and their combination into a single look-up table, here called
a {\tt FastKernel} ({\tt FK}) table, in such a way that the prediction
for a given hadronic observable can be obtained by a simple matrix
product between the respective {\tt FK} table and the initial scale
PDFs.

In this paper we present the {\tt APFELgrid} package, a public
implementation of the {\tt FastKernel} method in which the hard
partonic cross sections provided in an {\tt APPLgrid} look-up table
are combined with the DGLAP evolution kernels provided by the {\tt
  APFEL}
package~\cite{Bertone:2013vaa,Carrazza:2014gfa,Bertone:2015cwa,Bertone:2015lqa,Carrazza:2015dea,Bertone:2015gba}.

The outline of this paper is the following. In
Sect.~\ref{sec:FastKernel} we present the technical details of the
implementation of the {\tt FastKernel} method. Then, in
Sect.~\ref{sec:benchmark} we illustrate timings and performance
results of the {\tt APFELgrid} library and grids. Finally, in
Sect.~\ref{sec:conclusion} we summarise the results discussed in this
work.

@@

Hadron collider observables are typically computed in QCD by a double
convolution of parton densities with a hard scattering
cross-section. Consider for example the calculation of a general
cross-section $pp\to X$ with a set of PDFs $\{f\}$:
\begin{equation}\label{eq:hadconv}
  \sigma_{pp\to X} =
  \sum_{s}\sum_{p} \int dx_1\,dx_2\,
  \hat{\sigma}^{(p)(s)}\,\alpha_s^{p+p_{\rm LO}}(Q^2) \,F^{(s)}(x_1,x_2, Q^2)\,.
\end{equation}
where $Q^2$ is the typical hard scale of the process, the index $s$
sums over the active partonic subprocesses in the calculation, $p_{\rm
  LO}$ is the leading-order power of $\alpha_s$ for the process and
$\hat{\sigma}^{(p)(s)}$ is the N$^p$LO contribution to the
cross-section for the partonic subprocess scattering $s\to
X$. $F^{(s)}$ represents the subprocess parton density:
\begin{equation}\label{eq:APPLsubproc}
  F^{(s)}(x_1,x_2, Q^2) =\sum_{i,j} C^{(s)}_{ij} 
  f_i(x_{1},Q^2)f_j(x_{2},Q^2)\,,
\end{equation}
where the $C^{(s)}_{ij}$ matrix enumerates the combinations of PDFs
that contribute to the $s$-th subprocess.  The central observation of
tools such as {\tt APPLgrid} and {\tt FastNLO} is that the PDF and
$\alpha_S$ dependence may be factorised out of the convolution via
expansion over a set of interpolating functions, spanning $Q^2$ and
the two values of parton-$x$. For example one may represent the
subprocess PDFs and $\alpha_S$ in terms of the Lagrange basis
polynomials $\mathcal{I}_\tau(Q^2)$, $\mathcal{I}_\alpha(x_1)$ and
$\mathcal{I}_\beta(x_2)$ as:
\begin{equation}\label{eq:interpolation}
\begin{array}{c}
\displaystyle \alpha_s^{p+p_{\rm LO}}(Q^2) \,F^{(s)}(x_1,x_2, Q^2)
  =\\
\\
\displaystyle \sum_{\alpha,\beta,\tau} \alpha_s^{p+p_{\rm LO}}(Q_\tau^2)
  \,F^{(s)}_{\alpha\beta , \tau}
  \,\mathcal{I}_\tau(Q^2)\,\mathcal{I}_\alpha(x_1)
  \,\mathcal{I}_\beta(x_2)
\end{array}
\end{equation}
where we use the shorthand $F^{(s)}_{\alpha\beta ,\tau} =
F^{(s)}(x_\alpha, x_\beta,Q_\tau^2)$.  Using these expressions in the
double convolution of Eq.~(\ref{eq:hadconv}) one can finally obtain an
expression for the desired cross-section which depends upon the
subprocess PDFs only through a simple product:

\begin{equation} \label{eq:applconv}
  \sigma_{pp\to X} = \sum_p \sum_{s}\sum_{\alpha,\beta,\tau} 
  \alpha_s^{p+p_{\rm LO}}(Q^2_\tau)W_{\alpha\beta,\tau}^{(p)(s)} \, F_{\alpha\beta,\tau}^{(s)}\,.
\end{equation}
where:
\begin{equation}
  W_{\alpha\beta,\tau}^{(p)(s)} = \mathcal{I}_\tau(Q^2)\int dx_1\,dx_2\,
  \hat{\sigma}^{(p)(s)}\,\mathcal{I}_\alpha(x_1)
  \,\mathcal{I}_\beta(x_2)\,,
\end{equation}
consists of the convolution of the hard cross-section with the
interpolating polynomials. This information may be stored in a
precomputed look-up table.  The final expression for the cross-section
in Eq.~(\ref{eq:applconv}) therefore being considerably more
computationally efficient inside a fit than the direct performance of
the double convolution.

@@

A number of tools, $e.g.$ {\tt
  APFEL}~\cite{Bertone:2013vaa,Carrazza:2014gfa,Bertone:2015cwa,Bertone:2015lqa,Carrazza:2015dea,Bertone:2015gba},
{\tt HOPPET}~\cite{Salam:2008qg} and {\tt QCDNUM}~\cite{Botje:2010ay},
are available which perform PDF evolution via an analogous
procedure. {\tt APFEL} in particular, is able to provide the so-called
\textit{evolution operator}. The evolution operator, is computed
assuming that the DGLAP evolution can be encoded in a single operator
whose action on the intial scale PDFs proceduces the final scale
PDFs. Taking $Q_0$ as an initial scale and $Q_\tau$ as a final scale,
on the $x$-space grid one can write:
\begin{equation}\label{eq:fastPDFfinal_recalled}
  f_i(x_{\alpha},Q^2_\tau) = \sum_{k}
  \sum_\beta A^\tau_{\alpha\beta, ik}\;
  f_k(x_\beta,Q^2_0)\,, 
\end{equation}
where the operator $A$ is the solution of the DGLAP evolution equation
with baundary condition:
\begin{equation}
A^\tau_{\alpha\beta, ik} =
\delta_{\alpha\beta}\delta_{ik}\qquad\mbox{for}\qquad Q_\tau = Q_0\,.
\end{equation}

Given the evolution operator defined in
Eq.~(\ref{eq:fastPDFfinal_recalled}), we may replace the
(general-scale) PDFs used in the subprocess parton density
Eq.~(\ref{eq:APPLsubproc}) with their equivalent expressions like so:
\begin{equation}\label{eq:FK1}
\begin{array}{rcl}
F^{(s)}_{\alpha\beta,\tau} &=&  \displaystyle \sum_{i,j} \sum_{k,l}
                               \sum_{\delta,\gamma} C^{(s)}_{ij}
                               \left[  A^\tau_{\alpha\delta ik}\;
                               f_k(x_\delta,Q^2_0) A^\tau_{\beta\gamma
                               jl}\; f_l(x_\gamma,Q^2_0) \right]\;\;\;
  \\
\\
&=& \displaystyle \sum_{k,l}\sum_{\delta,\gamma}
\widetilde{C}^{(s),\tau}_{kl,\alpha\beta\gamma\delta}
f_k(x_\delta,Q^2_0) f_l(x_\gamma,Q^2_0)\,,
\end{array}
\end{equation}
where the object:
\begin{equation}
  \widetilde{C}^{(s),\tau}_{kl,\alpha\beta\gamma\delta} =
  \sum_{i,j} C^{(s)}_{ij} A^\tau_{\alpha\delta ik}
  A^\tau_{\beta\gamma jl}\,,
\end{equation}
combines the operations of subprocess density construction and PDF
evolution. Going further it is possible to then use the expression for
subprocess parton densities in Eq.~(\ref{eq:FK1}) in the full
cross-section calculation:
\begin{equation}
\sigma_{pp\to X} = \sum_{k,l}\sum_{\delta,\gamma}\sum_p
\sum_{s} \sum_{\alpha,\beta}
\sum_{\tau} 
\alpha_s^{p+p_{\rm LO}}(Q^2_\tau)W_{\alpha\beta,\tau}^{(p)(s)} \widetilde{C}^{(s),\tau}_{kl,\alpha\beta\gamma\delta}
f_k(x_\delta,Q^2_0) f_l(x_\gamma,Q^2_0)\,.
\end{equation}
Performing some further contractions it is possible to obtain an
extremely compact expression for the calculation of the cross section
in question, in terms of only the initial-scale PDFs and summing only
over the initial scale interpolating $x$-grid and the incoming parton
flavors:
\begin{equation}\label{eq:FK}
  \sigma _{pp\to X} = \sum_{k,l}\sum_{\delta,\gamma} 
  \widetilde{W}_{kl,\delta\gamma} \,f_k(x_\delta,Q^2_0) f_l(x_\gamma,Q^2_0),
\end{equation}
where the object:
\begin{equation}\label{eq:FKTable}
  \widetilde{W}_{kl,\delta\gamma} = \sum_p\sum_{s}\sum_{\alpha,\beta} \sum_{\tau}
\alpha_s^{p+p_{\rm LO}}(Q^2_\tau)  W_{\alpha\beta,\tau}^{(p)(s)} \widetilde{C}^{(s),\tau}_{kl,\alpha\beta\gamma\delta}
\end{equation}
is referred to here as an {\tt FK} table, and combines the information
stored in {\tt APPLgrid}-style interpolated weight grids with
analogously interpolated DGLAP evolution operators. This combination
enables for a maximally computationally-efficient expression for the
calculation of observables at hadron colliders under PDF variation.

@@

In this section we shall examine the performance differences between
the two expressions for cross-section prediction
Eq.~(\ref{eq:applconv}) ({\tt APPLgrid}) and Eq.~(\ref{eq:FK}) ({\tt
  FK}).
%
To provide a comprehensive benchmark, we consider here a wide range of
processes including LHC and Tevatron electroweak vector boson
production
measurements~\cite{Aaij:2012mda,Aaij:2012vn,Chatrchyan:2013mza,Chatrchyan:2013uja,Chatrchyan:2012xt,Aad:2013iua,Aad:2011fp,Aad:2011dm,Aaltonen:2010zza},
$t\bar{t}$ total
cross-sections~\cite{ATLAS:2012aa,ATLAS:2011xha,TheATLAScollaboration:2013dja,Chatrchyan:2013faa,Chatrchyan:2012bra,Chatrchyan:2012ria},
double-differential Drell-Yan
cross-sections~\cite{Chatrchyan:2013tia,CMS:2014jea} and inclusive jet
data~\cite{Chatrchyan:2012bja,Aad:2011fc,Aad:2013lpa,Abazov:2007jy}. Predictions
are performed over a wide range of kinematics, for a total of 52
source {\tt APPLgrid} files corresponding to the majority of available
LHC and Tevatron datasets of applicability to PDF determination.
While the source {\tt APPLgrid} files have varying grid densities in
$x$ and $Q^2$, for the purposes of comparison the corresponding {\tt
  FK} tables are produced consistently with 30 points in $x$, and at
an initial scale below the charm threshold, therefore with seven
active partonic species.

>>>>

\begin{abstract}
  We present a new software package designed to reduce the computational
  burden of collider measurements in Parton Distribution Function (PDF)
  fits. The {\tt APFELgrid} package converts interpolated weight
  tables provided by {\tt APPLgrid} files into a more efficient format
  for PDF fitting by the combination with PDF and
  $\alpha_s$ evolution factors provided by {\tt APFEL}. This
  combination significantly reduces the number of operations required
  to perform the calculation of hadronic observables in PDF fits and
  simplifies the structure of the calculation into a readily optimised
  scalar product. We demonstrate that our technique can lead to a
  substantial speed improvement when compared to existing methods
  without any reduction in numerical accuracy.
  % We present a new method conceived to ease the inclusion of
  % hadronic observables into parton distribution function (PDF)
  % fits. Our technique relies on the the same principles adopted by
  % the existing fast interfaces but implements a set of improvements
  % aimed at optimizing the performance in the context of PDF fits
  % where many iterations and many predictions for each iteration are
  % usually required. The main novelties are the precomputation of the
  % PDF and $\alpha_s$ evolution and the optimisation of the numerical
  % convolution between hard cross sections and PDFs. We demonstrate
  % that our technique can lead to a substantial speed improvement as
  % compared to the existing interfaces.
\end{abstract}

@@

Measurements at the Large Hadron Collider (LHC) have a unique capacity to shed light upon the internal dynamics 
of the proton and provide new constraints upon proton PDFs. However the inclusion of large collider datasets 
into PDF fits can provide a significant challenge due to the large computational footprint of performing accurate
theoretical predictions over the many iterations required in a fitting procedure. In order to make the fullest use of current 
and future LHC results, efficient strategies for the computation of these observables must therefore be employed.

The typical Monte Carlo software packages used to perform predictions for 
hadron-collider observables cannot be easily deployed in a PDF fit due the processing time 
required to obtain accurate results (usually of the order of a few hours or more per data
point). To overcome such limitations, the typical strategy
adopted for fast cross section prediction relies on the precomputation
of the partonic hard cross sections in such a way that the standard numerical
convolution with any set of PDFs can be reliably approximated by means of
interpolation techniques.

Such interpolation strategies are implementated in the {\tt
  APPLgrid}~\cite{Carli:2010rw} and {\tt
  FastNLO}~\cite{Wobisch:2011ij} projects. For the computation of the
hard cross sections, these packages rely on external codes to which
they are interfaced by means of a suite of functions allowing for the
filling of PDF- and $\alpha_s$-independent look-up tables of
cross section weights. Monte Carlo programs such as {\tt MCFM}~\cite{Campbell:2010ff}
and {\tt NLOJet++}~\cite{Nagy:2003tz} have been interfaced directly to
{\tt APPLgrid}/{\tt FastNLO} and more recently de\-di\-ca\-ted
interfaces to automated general-purpose event
generators have been developed. The {\tt
  aMCfast}~\cite{Bertone:2014zva} and {\tt
  MCgrid}~\cite{DelDebbio:2013kxa} codes can generate interpolation grids in {\tt APPLgrid}/{\tt FastNLO} format 
 by extracting the relevant information from the {\tt
  MadGraph5\_aMC@NLO}~\cite{Alwall:2014hca} and {\tt
  SHERPA}~\cite{Gleisberg:2008ta} event generators respectively.

While these tools have proven to be invaluable in the extraction
of parton densities, the volume of experimental data made available by
LHC collaborations for use in PDF fits is already stretching the
capabilities of the typical fitting technology. A typical
global PDF fit may now include thousands of hadronic data points for which
predictions have to be computed thousands of times during the
minimisation process. As a consequence performing these predictions
using the standard interpolating tools, $i.e.$
{\tt APPLgrid} and {\tt FastNLO}, starts to become prohibitively time-consuming.
For this reason a high-performance tool tailored specifically to the requirements of PDF analysis becomes increasingly
important.

The {\tt FastKernel} method was developed to
address this problem in the context of the NNPDF global analyses~\cite{Ball:2010de,Ball:2014uwa}. This method differs from the standard procedure
\`{a} la {\tt APPLgrid} or {\tt FastNLO} in that it maximises the
amount of information that is precomputed prior to fitting so as to
minimise the amount of operations required during the
fit. More specifically, the {\tt FastKernel} method relies on the
combination of precomputed hard cross sections with DGLAP evolution
kernels into a single look-up table, here called
a {\tt FastKernel} ({\tt FK}) table. In this way the prediction
for a given hadronic observable can be obtained by performing a simple matrix
product between the respective {\tt FK} table and PDFs evaluated directly at the fitting scale.

In this paper we present the {\tt APFELgrid} package, a public
implementation of the {\tt FastKernel} method in which the hard
partonic cross sections provided in an {\tt APPLgrid} look-up table
are combined with the DGLAP evolution kernels provided by the {\tt
  APFEL}
package~\cite{Bertone:2013vaa}.

This paper proceeds as follows. In
Sect.~\ref{sec:FastKernel} we present the technical details of the
implementation of the {\tt FastKernel} method. This is followed in
Sect.~\ref{sec:benchmark} by a performance benchmark of the
{\tt APFELgrid} library and resulting {\tt FK} tables. Finally, in
Sect.~\ref{sec:conclusion} we summarise the results discussed in this
work.

@@

Hadron collider observables are typically computed in QCD by means of a double
convolution of parton densities with a hard scattering
cross section. Consider for example the calculation of a general
cross section $pp\to X$ with a set of PDFs $\{f\}$:
\begin{equation}\label{eq:hadconv}
  \sigma_{pp\to X} =
  \sum_{s}\sum_{p} \int dx_1\,dx_2\,
  \hat{\sigma}^{(p)(s)}\,\alpha_s^{p+p_{\rm LO}}(Q^2) \,F^{(s)}(x_1,x_2, Q^2)\,,
\end{equation}
where $Q^2$ is the typical hard scale of the process, the index $s$
sums over the active partonic subprocesses in the calculation, $p$ sums over
the perturbative orders used in the expansion, $p_{\rm
  LO}$ is the leading-order power of $\alpha_s$ for the process and
$\hat{\sigma}^{(p)(s)}$ is the N$^p$LO contribution to the
cross section for the partonic subprocess scattering $(s)\to
X$. $F^{(s)}$ represents the subprocess parton density:
\begin{equation}\label{eq:APPLsubproc}
  F^{(s)}(x_1,x_2, Q^2) =\sum_{i,j} C^{(s)}_{ij} 
  f_i(x_{1},Q^2)f_j(x_{2},Q^2)\,,
\end{equation}
where the $C^{(s)}_{ij}$ matrix enumerates the combinations of PDFs
contributing to the $s$-th subprocess.  The central observation of
tools such as {\tt APPLgrid} and {\tt FastNLO} is that the PDF and
$\alpha_S$ dependence may be factorised out of the convolution via
expansion over a set of interpolating functions, spanning $Q^2$ and
the two values of parton-$x$. For example one may represent the
subprocess PDFs and $\alpha_S$ in terms of Lagrange basis
polynomials $\mathcal{I}_\tau(Q^2)$, $\mathcal{I}_\alpha(x_1)$ and
$\mathcal{I}_\beta(x_2)$ as:
\begin{equation}\label{eq:interpolation}
\begin{array}{c}
\displaystyle \alpha_s^{p+p_{\rm LO}}(Q^2) \,F^{(s)}(x_1,x_2, Q^2)
  =\\
\\
\displaystyle \sum_{\alpha,\beta,\tau} \alpha_s^{p+p_{\rm LO}}(Q_\tau^2)
  \,F^{(s)}_{\alpha\beta , \tau}
  \,\mathcal{I}_\tau(Q^2)\,\mathcal{I}_\alpha(x_1)
  \,\mathcal{I}_\beta(x_2),
\end{array}
\end{equation}
where we use the shorthand $F^{(s)}_{\alpha\beta ,\tau} =
F^{(s)}(x_\alpha, x_\beta,Q_\tau^2)$.  Using these expressions in the
double convolution of Eq.~(\ref{eq:hadconv}) one can finally obtain an
expression for the desired cross section which depends upon the
subprocess PDFs only through a simple product:

\begin{equation} \label{eq:applconv}
  \sigma_{pp\to X} = \sum_p \sum_{s}\sum_{\alpha,\beta,\tau} 
  \alpha_s^{p+p_{\rm LO}}(Q^2_\tau)W_{\alpha\beta,\tau}^{(p)(s)} \, F_{\alpha\beta,\tau}^{(s)}\,,
\end{equation}
where
\begin{equation}
  W_{\alpha\beta,\tau}^{(p)(s)} = \mathcal{I}_\tau(Q^2)\int dx_1\,dx_2\,
  \hat{\sigma}^{(p)(s)}\,\mathcal{I}_\alpha(x_1)
  \,\mathcal{I}_\beta(x_2)\,,
\end{equation}
consists of the convolution of the hard cross section with the
interpolating polynomials. This information may be stored in a
precomputed look-up table.  The final expression for the cross section
in Eq.~(\ref{eq:applconv}) is therefore a considerably simpler task to perform
inside a fit than the direct evaluation of the double convolution.

@@

A number of tools, $e.g.$ {\tt
  APFEL},
{\tt HOPPET}~\cite{Salam:2008qg} and {\tt QCDNUM}~\cite{Botje:2010ay},
are available which perform PDF evolution via an analogous
procedure. {\tt APFEL} in particular, is able to provide the so-called
\textit{evolution operator}. The evolution operator, is computed
assuming that the DGLAP evolution can be encoded in a single operator
whose action on the intial scale PDFs proceduces the final scale
PDFs. Taking $Q_0$ as an initial scale and $Q_\tau$ as a final scale,
on the $x$-space grid one can write:
\begin{equation}\label{eq:fastPDFfinal_recalled}
  f_i(x_{\alpha},Q^2_\tau) = \sum_{k}
  \sum_\beta A^\tau_{\alpha\beta, ik}\;
  f_k(x_\beta,Q^2_0)\,, 
\end{equation}
where the operator $A$ is the solution of the DGLAP evolution equation
with baundary condition:
\begin{equation}
A^\tau_{\alpha\beta, ik} =
\delta_{\alpha\beta}\delta_{ik}\qquad\mbox{for}\qquad Q_\tau = Q_0\,.
\end{equation}

Given the evolution operator defined in
Eq.~(\ref{eq:fastPDFfinal_recalled}), we may replace the
(general-scale) PDFs used in the subprocess parton density
Eq.~(\ref{eq:APPLsubproc}) with their equivalent expressions like so:
\begin{equation}\label{eq:FK1}
\begin{array}{rcl}
F^{(s)}_{\alpha\beta,\tau} &=&  \displaystyle \sum_{i,j} \sum_{k,l}
                               \sum_{\delta,\gamma} C^{(s)}_{ij}
                               \left[  A^\tau_{\alpha\delta ik}\;
                               f_k(x_\delta,Q^2_0) A^\tau_{\beta\gamma
                               jl}\; f_l(x_\gamma,Q^2_0) \right]\;\;\;
  \\
\\
&=& \displaystyle \sum_{k,l}\sum_{\delta,\gamma}
\widetilde{C}^{(s),\tau}_{kl,\alpha\beta\gamma\delta}
f_k(x_\delta,Q^2_0) f_l(x_\gamma,Q^2_0)\,,
\end{array}
\end{equation}
where the object:
\begin{equation}
  \widetilde{C}^{(s),\tau}_{kl,\alpha\beta\gamma\delta} =
  \sum_{i,j} C^{(s)}_{ij} A^\tau_{\alpha\delta ik}
  A^\tau_{\beta\gamma jl}\,,
\end{equation}
combines the operations of subprocess density construction and PDF
evolution. Going further it is possible to then use the expression for
subprocess parton densities in Eq.~(\ref{eq:FK1}) in the full
cross section calculation:
\begin{equation}
\sigma_{pp\to X} = \sum_{k,l}\sum_{\delta,\gamma}\sum_p
\sum_{s} \sum_{\alpha,\beta}
\sum_{\tau} 
\alpha_s^{p+p_{\rm LO}}(Q^2_\tau)W_{\alpha\beta,\tau}^{(p)(s)} \widetilde{C}^{(s),\tau}_{kl,\alpha\beta\gamma\delta}
f_k(x_\delta,Q^2_0) f_l(x_\gamma,Q^2_0)\,.
\end{equation}
Performing some further contractions it is possible to obtain an
extremely compact expression for the calculation of the cross section
in question, in terms of only the initial-scale PDFs and summing only
over the initial scale interpolating $x$-grid and the incoming parton
flavors:
\begin{equation}\label{eq:FK}
  \sigma _{pp\to X} = \sum_{k,l}\sum_{\delta,\gamma} 
  \widetilde{W}_{kl,\delta\gamma} \,f_k(x_\delta,Q^2_0) f_l(x_\gamma,Q^2_0),
\end{equation}
where the object:
\begin{equation}\label{eq:FKTable}
  \widetilde{W}_{kl,\delta\gamma} = \sum_p\sum_{s}\sum_{\alpha,\beta} \sum_{\tau}
\alpha_s^{p+p_{\rm LO}}(Q^2_\tau)  W_{\alpha\beta,\tau}^{(p)(s)} \widetilde{C}^{(s),\tau}_{kl,\alpha\beta\gamma\delta}
\end{equation}
is referred to here as an {\tt FK} table, and combines the information
stored in {\tt APPLgrid}-style interpolated weight grids with
analogously interpolated DGLAP evolution operators. This combination
enables for a maximally computationally-efficient expression for the
calculation of observables at hadron colliders under PDF variation.

@@

In this section we shall examine the performance differences between
the two expressions for cross section prediction
Eq.~(\ref{eq:applconv}) ({\tt APPLgrid}) and Eq.~(\ref{eq:FK}) ({\tt
  FK}).
%
To provide a comprehensive benchmark, we consider here a wide range of
processes including LHC and Tevatron electroweak vector boson
production
measurements~\cite{Aaij:2012mda,Aaij:2012vn,Chatrchyan:2013mza,Chatrchyan:2013uja,Chatrchyan:2012xt,Aad:2013iua,Aad:2011fp,Aad:2011dm,Aaltonen:2010zza},
$t\bar{t}$ total
cross sections~\cite{ATLAS:2012aa,ATLAS:2011xha,TheATLAScollaboration:2013dja,Chatrchyan:2013faa,Chatrchyan:2012bra,Chatrchyan:2012ria},
double-differential Drell-Yan
cross sections~\cite{Chatrchyan:2013tia,CMS:2014jea} and inclusive jet
data~\cite{Chatrchyan:2012bja,Aad:2011fc,Aad:2013lpa,Abazov:2007jy}. Predictions
are performed over a wide range of kinematics, for a total of 52
source {\tt APPLgrid} files corresponding to the majority of available
LHC and Tevatron datasets of applicability to PDF determination.
While the source {\tt APPLgrid} files have varying grid densities in
$x$ and $Q^2$, for the purposes of comparison the corresponding {\tt
  FK} tables are produced consistently with 30 points in $x$, and at
an initial scale below the charm threshold, therefore with seven
active partonic species.
